Substructure Boosts to Dark Matter Annihilation from Sommerfeld Enhancement

### 2009-03-03 tex/somm.tex
Removed old figures; fixed small integration limit error
Fixed some spelling mistakes, made figures nicer

\begin{abstract}
I discuss the implications of the recently introduced Sommerfeld
enhancement of the dark matter annihilation cross section for the
detection of subhalos in the Galactic halo.  In addition to the boost
to the dark matter annihilation cross section from the high densities
of these subhalos with respect to the main halo, an additional boost
due to Sommerfeld enhancement results from the fact that they are
kinematically colder than the Galactic halo. If we further believe the
generic prediction of \CDM\ that in each subhalo there is an abundance
of substructure which is approximately self-similar to that of the
Galactic halo, then additional boosts coming from the density
enhancements of these small substructures and their small velocity
dispersions enhance the dark matter annihilation cross section even
further. I show that very large boost factors ($10^5$ -- $10^9$) are
obtained in a large class of models. The implications of these boost
factors for the detection of dark matter annihilation from dwarf
Speroidal galaxies in the Galactic halo are such that they outshine
the background gamma-ray flux and should be detectable
by \Fermi.
\end{abstract}

@@

Of course, the largest obstacle currently to modeling the PAMELA/ATIC
signal with \DM\ annihilation is the very large boost factors ($10^4$
-- $10^5$) with respect to the fiducial $\langle \sigma v
\rangle_{\mathrm{ann}} \approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$
needed to lead to an observable annihilation flux. The boost factor
from the clumpiness of the dark matter halo is expected to be small
\cite{2008A&A...479..427L,2008Natur.454..735D,2008ApJ...686..262K,2008Natur.456...73S},
typically of the order 1 -- 10. Therefore, a different mechanism must
be reponsible for the large boosts necessary. One such mechanism that
has received much attention lately is the Sommerfeld enhancement
\cite{sommerfeld31a,2003PhRvD..67g5014H,2004PhRvL..92c1303H,2005PhRvD..71f3528H,2005PhRvD..71a5007H,2006PhRvD..73e5004H,2008NuPhB.800..204C,2008JHEP...07..058M,2008arXiv0812.0559M,2008arXiv0812.0360L}. The
Sommerfeld enhancement is a generic effect present whenever there is
an attractive force acting between the dark matter particles, \eg, due
to a Yukawa interaction or a gauge interaction through vector bosons
\cite{ArkaniHamed:2008qn}. The main effect of this attractive force is
to enhance the annihilation cross section $\sigma v$ with a factor
proportional to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as
``$1/v$'' enhancement. There are regions in the parameter space made
up of the dark matter mass \mdm, the force carrier mass \mv, and the
force ``fine-structure constant'' $\alpha \equiv$ coupling$^2$/$4 \pi$
in which the enhancement is near a resonance, where it approximately
behaves as $\beta^{-2}$. However, the enhancement levels off at a
certain $\beta$ due to the finite range of the attractive force ---
large ($\alpha \sim\!10^{-1}$ -- $10^{-3}$) attractive forces with
infinite range lead to a burst of \DM\ annihilation in the first \DM\
halos formed at $z$ $\sim$ 100 -- 200 which would lead to observable
effects incompatible with measurements of the diffuse extragalactic
gamma-ray background today and the cosmic microwave background
\cite{2008arXiv0810.3233K}. Since the typical velocities of dark
matter particles in the Galactic halo today are of the order of the
velocity dispersion of the halo, which is non-relativistic ($\beta\sim
10^{-3}$), the cross section of dark matter particles annihilating in
the halo could be significantly enhanced by this Sommerfeld
enhancement.

@@

In this paper we quantify the magnitude of this new substructure boost
and discuss its implications for the detection of dark matter
annihilation from subhalos in the Galactic halos. As we will show, the
total combined boost from Sommerfeld enhancement and substructure can
reach very large values (10$^5$ -- 10$^9$), and these boosts are only
weakly dependent on the assumptions that we have to make about the
abundance of substructure and the \DM\ density profile. We use these
total boosts together with published \DM\ density profiles of some of
the dwarf Spheroidal satellite galaxies of the Milky Way to predict
the total flux of gamma-rays from \DM\ annihilation that would be
detectable by the Fermi Gamma-ray Space Telescope (\Fermi) and find
that in most of the models that posit the Sommerfeld enhancement the
\DM\ annihilation signal from the dwarf Spheroidals shines above the
diffuse extragalactic gamma-ray background, which is the only
background signal at the high Galactic lattitudes of these satellite
galaxies. Finally, we also consider the prospects for detecting
smaller subhalos of the Galaxy by computing annihilation fluxes of
subhalos in the \VL\ II simulation. We will see that whether these
subhalos will be detectable in the near future is highly dependent on
the exact details of the Sommerfeld enhancement.

@@

\begin{figure*}[t]
\includegraphics[clip=]{avg-3_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-4_129.eps}\\%%BoundingBox: 190 291 418 474
\includegraphics[clip=]{avg-5_129.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-6_129.eps}%%BoundingBox: 190 291 418 474
\caption{Sommerfeld enhancement factor as a function of the particle
physics parameters assuming a one-particle Maxwell-Boltzman velocity
distribution with velocity dispersion \sigv\ for four different orders
of magnitude of the velocity dispersion (all velocities are relative
to the speed of light).}%
\label{fig:sommavg}%
\end{figure*}

@@

\section{Substructure boosts from Sommerfeld enhancement}
The photon flux from \DM\ annihilation from a solid angle
$\Delta\Omega$ along a given line-of-sight (\los) is given by
\begin{equation}\label{eq:flux}
\begin{split}
\frac{\dd \Ngamma}{\dd A \,\dd t}  = &
\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
\Ngammai}{\dd E} \,\frac{\sigmaannv_i}{\mdm^2}\,
\somm\left(\mv/\mdm,\alpha,\sigv\right)\,\\
& \times \int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd l\,
\rho^2(l)\, .
\end{split}
\end{equation}
In this \Eth\ is the treshold energy of the detector, $i$ denotes
(possibly) different final states of the \DM\ annihilation, $\frac{\dd
\Ngammai}{\dd E}$ is the spectrum of the annihilation to that state,
$\sigmaannv_i$ is the annihilation cross section to state $i$ (without
the Sommerfeld enhancement), $\somm\left(\mv/\mdm,\alpha,\beta\right)$
is the Sommerfeld enhancement which we have factored out of the cross
section, and $\rho(l)$ is the \DM\ density along the line-of-sight. We
emphasize that this framework also allows us to consider different
final annihilation products, such as the electrons and positrons
supposedly seen by PAMELA/ATIC.

@@

For the annihilation cross section without taking account of any
Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
3 \times 10^{-26} $ cm$^3$ s$^{-1}$, and we choose a fiducial mass for
the \DM\ particle of \mdm\ = 700 GeV. The energy spectrum of the \DM\
annihilation depends on the model used to describe the \DM\ and its
interactions and it is outside of the scope of this paper to provide a
detailed treatment of this for different \DM\ models. Therefore, we
will adopt an annihilation spectrum for \DM\ annihilation into photons
appropriate for the newer models that were proposed to explain the
PAMELA/ATIC results. Specifically, we use a gamma ray spectrum from a
cascade annihilation model. In these models the \DM\ annihilates into
a new light degree of freedom which then in turn decays into leptons
and photons. The injection spectrum per \DM\ annihilation is given by
\cite{2009arXiv0901.2926M}
\begin{equation}\label{eq:annspectrum}
\begin{split}
\frac{\dd \Ngamma}{\dd E} = &\frac{\alpha_{\mbox{{\footnotesize EM}}}}{\pi} \frac{1+(1-E/\mdm)^2}{E}\\
&\times\left\{-1 + \ln\left(4(1-E/\mdm)\right) -2\ln\left(\frac{m_l}{\mdm}\right)\right\}\, ,
\end{split}
\end{equation}
in which $\alpha_{\mbox{{\footnotesize EM}}}$ is the electromagnetic
coupling constant and $m_l$ is the mass of the leptons created in the
process $\chi\chi \rightarrow l^+l^-\gamma$. Integrating this spectrum
using an energy range appropriate for \Fermi\ we find that
\begin{equation} 
\int_{4 \mbox{{\footnotesize GeV}}}^{250 \mbox{{\footnotesize GeV}}}\dd E\, \frac{\dd
\Ngamma}{\dd E} \approx 0.5\, .
\end{equation}
We find a similar number if we integrate the spectrum due to one
cascade step. It is interesting to note that these photon spectra have
fewer photons in them, about an order of magnitude less, than the
annihilation spectra due to neutralino annihilation which are often
used \cite{1998APh.....9..137B,2004PhRvD..70j3529F}. Of course, since
the energy spectrum contributes a multiplicative factor to the \DM\
annihilation flux, the effect of different annihilation spectra can be
incorporated by simply multiplying our results with the appropriate
factor.

@@

In the previous section we calculated $\lumsmoothc$ for the different
\DM\ density profiles and found in all cases that
\begin{equation}
\lumsmoothc(m) \equiv \lumsmoothc(\rho(m),r(m))\, ,
\end{equation}
for a characteristic density $\rho$ and radius $r$, the exact
definition of which depends on the specific \DM\ density profile
used. In order to find the mass dependence of $\lumsmoothc$, we use
the description of the \DM\ density profile in terms of virial mass
\Mvir\ and concentration \conc. We then make use of the fact that many
simulations have found a relation between the concentration of halos
and their mass
\cite{2001MNRAS.321..559B,2001ApJ...554..114E,2008MNRAS.387..536G,2007MNRAS.381.1450N},
such that everything depends only on the mass. This relation is well
established over a large range in halo masses, and we will adopt the
particular relationship found in Ref.~\cite{2007MNRAS.378...55M}
\begin{equation}\label{eq:concscaling}
\conc = 10.5 \, \left(\frac{\Mvir}{10^{12}\, \Msol}\right)^{-0.11}\, ,
\end{equation}
in good agreement with the results of other, similar analyses
\cite{2001MNRAS.321..559B,2005MNRAS.357..387K}, although the
normalization of the concentration-mass relation is about 15\% lower
than the original relation. However, the concentration of subhalos is
generally larger than the concentration of field halos of the same
mass as a result of tidal mass-loss
\cite{1998MNRAS.300..146G,2008ApJ...673..226P,2004MNRAS.355..794H,2005ApJ...635..931B,2004ApJ...608..663K}. This
is a small effect, of order 10\%, and it does not play a large role in
the results we obtain below. This concentration-halo mass
concentration has not just been establised in numerical simulations:
statistical weak lensing analysis of luminous red galaxies and
clusters from the Sloan Digital Sky Survey (\SDSS) has shown that the
halo profiles of these objects are consistent with the \NFW\ profile
and mass-concentration relation given above
\cite{2008JCAP...08..006M}.

@@

However, the Sommerfeld enhancement has not saturated in all regions
of parameter space at this mass scale, and for smaller values of
\mv/\mdm and especially in resonance regions, the reduced boost can
reach high values, signalling a large extra boost due to the
kinematically coldness of the substructure.  For small values of
\mv/\mdm\ we see that we can get boosts that are an order of magnitude
to two orders of magnitude larger than what we would expect from
naively combining the Sommerfeld enhancement due to the velocity
dispersion of the subhalo with the boost from density enhancements
alone. This can be easily understood as being due to the fact that the
Sommerfeld enhancement has not saturated yet at the mass scale of the
top-level subhalo, such that it can be boosted up to it saturation
point by the contribution from the kinematically colder substructure.

@@

It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
predicted to be coming from the \dSphs\ is very large in a large class
of models. We have to stress here that our predicted fluxes are all
for the particle physics specific parameters given in Section
\ref{sec:partphys}, \ie, \sigmaannv\ = $3 \times 10^{-26}$ cm$^3$
s$^{-1}$, \mdm\ = 700 GeV and a particular form for the \DM\
annihilation spectrum. To predict the gamma-ray flux for a particular
model with different values of these particle physics parameters one
is required to calculate the particle physics parameter specific
factor to the \DM\ annihilation flux and appropriately rescale the
values given here. In order to compare these signals to the expected
background flux we will only consider contributions to the background
flux of extragalactic gamma rays. Although there are also
contributions to the background flux from pion-decay, inverse Compton
scattering, and Bremsstrahlung, at the high Galactic lattitudes at
which the dwarf Spheroidals are found these are all of the same order
or smaller than the extragalactic background
\cite{2004ApJ...613..962S}. We have integrated the extragalactic gamma
ray flux as measured by EGRET over the energy range between 4 GeV --
250 GeV and multiplied it with the solid angle that the \dSphs\ span
on the sky. The extragalactic gamma ray flux is given by a power law
\cite{1998ApJ...494..523S}
\begin{equation}
\frac{\dd \Ngamma}{\dd E \,\dd A \, \dd t\,\dd \Omega} = k\,\left(\frac{E}{E_0}\right)^{-\alpha}\, ,
\end{equation}
where $k = 7.3 \times 10^{-6}$ cm$^{-2}$ s$^{-1}$ sr$^{-1}$
GeV$^{-1}$, $\alpha$ = 2.10, and $E_0 = 0.45$ GeV. We have included
the background flux of extragalactic photons in Figure
\ref{fig:dsph}. This background flux is in all cases of the order of
10$^{-9}$ -- 10$^{-10}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
annihilation boosted by substructure and Sommerfeld enhancement
outshine this background in most of the parameter space, and by a few
orders of magnitude in most models. Given that the flux above 100 MeV
over a 55 days time-span needed to separate a point source of \DM\
annihilation from the background by \Fermi\ to give a 5$\sigma$
detection is about 10$^{-8}$ photons cm$^{-1}$ s$^{-1}$ and that the
necessary fluxes scale roughly as $t_{\mbox{{\footnotesize
exp}}}^{-1/2}$ \cite{2008JCAP...07..013B}, the \DM\ annihilation of
these \dSphs\ should be detectable by \Fermi\ in many of the models
parameterized by ($\mv/\mdm,\alpha$). The Ursa Major II \dSph\ is the
only \dSph\ for which the \DM\ annihilation dominates the background
in all the regions of parameter space. This shows that the newly
discovered, most-dark-matter-dominated satellite galaxies are possibly
the best targets for the indirect detection of \DM. We conclude that
the prospects for detecting \DM\ annihilation from the \dSphs\ with
\Fermi\ are very good in models that include a significant Sommerfeld
enhancement.

@@

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_2.eps}
\caption{Left: Histogram of the predicted gamma-ray flux coming from
indivual subhalos contained within the tidal radius of the main halo
in \VL\ II for three random observer placed at a distance of 8.5 kpc
from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
$\alpha = 10^{-3}$. Right: Histogram of the total boosts of the
subhalos for the same values of \mv/\mdm\ and $\alpha$.}%\
\label{fig:VL23_2}%
\end{figure}

@@

We only cover some typical cases for the Sommerfeld enhancement here
which suffices to show the general trend. We consider the particle
physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
smallest Sommerfeld enahcements in Figures \ref{fig:VL23_2} and
\ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
left panel of Figure \ref{fig:VL23_2} shows the distribution of
gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
right panel shows the distribution of the total boost factors. As can
be seen, the total boost factors are quite small: the typical boost
factor is only about 30 with the highest boost factors being about
100. The total gamma-ray flux coming from subhalos is then also quite
small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
the total gamma-ray flux is shown for 1,000 random observers at a
distance of 8.5 kpc from the galactic center. At most observation
locations this total flux is about 10$^{-7}$ cm$^2$ s$^{-1}$, with the
high-end tail not extending further than about half an order of
magnitude from this. For comparison the dashed histogram in Figure
\ref{fig:VL23_3} shows the same distribution of total gamma-ray flux
for different random observers in the absence of any Sommerfeld
enhancement. This confirms that the total gamma-ray flux with
Sommerfeld enhancement is about an order of magnitude larger than
without Sommerfeld enhancement.

@@

{\bf Acknowledgements:} It is a pleasure to thank Ilias Cholis, Mulin
Ding, David Hogg, and Neal Weiner for comments and assistance. This
research was partially supported by NASA (grant NNX08AJ48G).

>>>>

\begin{abstract}
I discuss the implications of the recently introduced Sommerfeld
enhancement of the dark matter annihilation cross section for the
detection of subhalos in the Galactic halo.  In addition to the boost
to the dark matter annihilation cross section from the high densities
of these subhalos with respect to the main halo, an additional boost
due to Sommerfeld enhancement results from the fact that they are
kinematically colder than the Galactic halo. If we further believe the
generic prediction of \CDM\ that in each subhalo there is an abundance
of substructure which is approximately self-similar to that of the
Galactic halo, then additional boosts coming from the density
enhancements of these small substructures and their small velocity
dispersions enhance the dark matter annihilation cross section even
further. I show that very large boost factors ($10^5$ -- $10^9$) are
obtained in a large class of models. The implications of these boost
factors for the detection of dark matter annihilation from dwarf
Spheroidal galaxies in the Galactic halo are such that they outshine
the background gamma-ray flux and should be detectable
by \Fermi.
\end{abstract}

@@

Of course, the largest obstacle currently to modeling the PAMELA/ATIC
signal with \DM\ annihilation is the very large boost factors ($10^4$
-- $10^5$) with respect to the fiducial $\langle \sigma v
\rangle_{\mathrm{ann}} \approx 3 \times 10^{-26} $ cm$^3$ s$^{-1}$
needed to lead to an observable annihilation flux. The boost factor
from the clumpiness of the dark matter halo is expected to be small
\cite{2008A&A...479..427L,2008Natur.454..735D,2008ApJ...686..262K,2008Natur.456...73S},
typically of the order 1 -- 10. Therefore, a different mechanism must
be responsible for the large boosts necessary. One such mechanism that
has received much attention lately is the Sommerfeld enhancement
\cite{sommerfeld31a,2003PhRvD..67g5014H,2004PhRvL..92c1303H,2005PhRvD..71f3528H,2005PhRvD..71a5007H,2006PhRvD..73e5004H,2008NuPhB.800..204C,2008JHEP...07..058M,2008arXiv0812.0559M,2008arXiv0812.0360L}. The
Sommerfeld enhancement is a generic effect present whenever there is
an attractive force acting between the dark matter particles, \eg, due
to a Yukawa interaction or a gauge interaction through vector bosons
\cite{ArkaniHamed:2008qn}. The main effect of this attractive force is
to enhance the annihilation cross section $\sigma v$ with a factor
proportional to $\beta^{-1}$ ($\beta \equiv v/c$), this is known as
``$1/v$'' enhancement. There are regions in the parameter space made
up of the dark matter mass \mdm, the force carrier mass \mv, and the
force ``fine-structure constant'' $\alpha \equiv$ coupling$^2$/$4 \pi$
in which the enhancement is near a resonance, where it approximately
behaves as $\beta^{-2}$. However, the enhancement levels off at a
certain $\beta$ due to the finite range of the attractive force ---
large ($\alpha \sim\!10^{-1}$ -- $10^{-3}$) attractive forces with
infinite range lead to a burst of \DM\ annihilation in the first \DM\
halos formed at $z$ $\sim$ 100 -- 200 which would lead to observable
effects incompatible with measurements of the diffuse extragalactic
gamma-ray background today and the cosmic microwave background
\cite{2008arXiv0810.3233K}. Since the typical velocities of dark
matter particles in the Galactic halo today are of the order of the
velocity dispersion of the halo, which is non-relativistic ($\beta\sim
10^{-3}$), the cross section of dark matter particles annihilating in
the halo could be significantly enhanced by this Sommerfeld
enhancement.

@@

In this paper we quantify the magnitude of this new substructure boost
and discuss its implications for the detection of dark matter
annihilation from subhalos in the Galactic halos. As we will show, the
total combined boost from Sommerfeld enhancement and substructure can
reach very large values (10$^5$ -- 10$^9$), and these boosts are only
weakly dependent on the assumptions that we have to make about the
abundance of substructure and the \DM\ density profile. We use these
total boosts together with published \DM\ density profiles of some of
the dwarf Spheroidal satellite galaxies of the Milky Way to predict
the total flux of gamma-rays from \DM\ annihilation that would be
detectable by the Fermi Gamma-ray Space Telescope (\Fermi) and find
that in most of the models that posit the Sommerfeld enhancement the
\DM\ annihilation signal from the dwarf Spheroidals shines above the
diffuse extragalactic gamma-ray background, which is the only
background signal at the high Galactic latitudes of these satellite
galaxies. Finally, we also consider the prospects for detecting
smaller subhalos of the Galaxy by computing annihilation fluxes of
subhalos in the \VL\ II simulation. We will see that whether these
subhalos will be detectable in the near future is highly dependent on
the exact details of the Sommerfeld enhancement.

@@

\begin{figure*}[t]
\includegraphics[clip=]{avg-3_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-4_129.eps}\\%%BoundingBox: 190 291 418 474
\includegraphics[clip=]{avg-5_129.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{avg-6_129.eps}%%BoundingBox: 190 291 418 474
\caption{Sommerfeld enhancement factor as a function of the particle
physics parameters assuming a one-particle Maxwell-Boltzmann velocity
distribution with velocity dispersion \sigv\ for four different orders
of magnitude of the velocity dispersion (all velocities are relative
to the speed of light).}%
\label{fig:sommavg}%
\end{figure*}

@@

\section{Substructure boosts from Sommerfeld enhancement}
The photon flux from \DM\ annihilation from a solid angle
$\Delta\Omega$ along a given line-of-sight (\los) is given by
\begin{equation}\label{eq:flux}
\begin{split}
\frac{\dd \Ngamma}{\dd A \,\dd t}  = &
\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
\Ngammai}{\dd E} \,\frac{\sigmaannv_i}{\mdm^2}\,
\somm\left(\mv/\mdm,\alpha,\sigv\right)\,\\
& \times \int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd l\,
\rho^2(l)\, .
\end{split}
\end{equation}
In this \Eth\ is the threshold energy of the detector, $i$ denotes
(possibly) different final states of the \DM\ annihilation, $\frac{\dd
\Ngammai}{\dd E}$ is the spectrum of the annihilation to that state,
$\sigmaannv_i$ is the annihilation cross section to state $i$ (without
the Sommerfeld enhancement), $\somm\left(\mv/\mdm,\alpha,\beta\right)$
is the Sommerfeld enhancement which we have factored out of the cross
section, and $\rho(l)$ is the \DM\ density along the line-of-sight. We
emphasize that this framework also allows us to consider different
final annihilation products, such as the electrons and positrons
supposedly seen by PAMELA/ATIC.

@@

For the annihilation cross section without taking account of any
Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
3 \times 10^{-26} $ cm$^3$ s$^{-1}$, and we choose a fiducial mass for
the \DM\ particle of \mdm\ = 700 GeV. The energy spectrum of the \DM\
annihilation depends on the model used to describe the \DM\ and its
interactions and it is outside of the scope of this paper to provide a
detailed treatment of this for different \DM\ models. Therefore, we
will adopt an annihilation spectrum for \DM\ annihilation into photons
appropriate for the newer models that were proposed to explain the
PAMELA/ATIC results. Specifically, we use a gamma ray spectrum from a
cascade annihilation model. In these models the \DM\ annihilates into
a new light degree of freedom which then in turn decays into leptons
and photons. The injection spectrum per \DM\ annihilation is given by
\cite{2009arXiv0901.2926M}
\begin{equation}\label{eq:annspectrum}
\begin{split}
\frac{\dd \Ngamma}{\dd E} = &\frac{\alpha_{\mbox{{\footnotesize EM}}}}{\pi} \frac{1+(1-E/\mdm)^2}{E}\\
&\times\left\{-1 + \ln\left(4(1-E/\mdm)\right) -2\ln\left(\frac{m_l}{\mdm}\right)\right\}\, ,
\end{split}
\end{equation}
in which $\alpha_{\mbox{{\footnotesize EM}}}$ is the electromagnetic
coupling constant and $m_l$ is the mass of the leptons created in the
process $\chi\chi \rightarrow l^+l^-\gamma$. Integrating this spectrum
using an energy range appropriate for \Fermi\ we find that
\begin{equation} 
\int_{4 \mbox{ {\footnotesize GeV}}}^{250 \mbox{ {\footnotesize GeV}}}\dd E\, \frac{\dd
\Ngamma}{\dd E} \approx 0.5\, .
\end{equation}
We find a similar number if we integrate the spectrum due to one
cascade step. It is interesting to note that these photon spectra have
fewer photons in them, about an order of magnitude less, than the
annihilation spectra due to neutralino annihilation which are often
used \cite{1998APh.....9..137B,2004PhRvD..70j3529F}. Of course, since
the energy spectrum contributes a multiplicative factor to the \DM\
annihilation flux, the effect of different annihilation spectra can be
incorporated by simply multiplying our results with the appropriate
factor.

@@

In the previous section we calculated $\lumsmoothc$ for the different
\DM\ density profiles and found in all cases that
\begin{equation}
\lumsmoothc(m) \equiv \lumsmoothc(\rho(m),r(m))\, ,
\end{equation}
for a characteristic density $\rho$ and radius $r$, the exact
definition of which depends on the specific \DM\ density profile
used. In order to find the mass dependence of $\lumsmoothc$, we use
the description of the \DM\ density profile in terms of virial mass
\Mvir\ and concentration \conc. We then make use of the fact that many
simulations have found a relation between the concentration of halos
and their mass
\cite{2001MNRAS.321..559B,2001ApJ...554..114E,2008MNRAS.387..536G,2007MNRAS.381.1450N},
such that everything depends only on the mass. This relation is well
established over a large range in halo masses, and we will adopt the
particular relationship found in Ref.~\cite{2007MNRAS.378...55M}
\begin{equation}\label{eq:concscaling}
\conc = 10.5 \, \left(\frac{\Mvir}{10^{12}\, \Msol}\right)^{-0.11}\, ,
\end{equation}
in good agreement with the results of other, similar analyses
\cite{2001MNRAS.321..559B,2005MNRAS.357..387K}, although the
normalization of the concentration-mass relation is about 15\% lower
than the original relation. However, the concentration of subhalos is
generally larger than the concentration of field halos of the same
mass as a result of tidal mass-loss
\cite{1998MNRAS.300..146G,2008ApJ...673..226P,2004MNRAS.355..794H,2005ApJ...635..931B,2004ApJ...608..663K}. This
is a small effect, of order 10\%, and it does not play a large role in
the results we obtain below. This concentration-halo mass
concentration has not just been established in numerical simulations:
statistical weak lensing analysis of luminous red galaxies and
clusters from the Sloan Digital Sky Survey (\SDSS) has shown that the
halo profiles of these objects are consistent with the \NFW\ profile
and mass-concentration relation given above
\cite{2008JCAP...08..006M}.

@@

However, the Sommerfeld enhancement has not saturated in all regions
of parameter space at this mass scale, and for smaller values of
\mv/\mdm and especially in resonance regions, the reduced boost can
reach high values, signaling a large extra boost due to the
kinematically coldness of the substructure.  For small values of
\mv/\mdm\ we see that we can get boosts that are an order of magnitude
to two orders of magnitude larger than what we would expect from
naively combining the Sommerfeld enhancement due to the velocity
dispersion of the subhalo with the boost from density enhancements
alone. This can be easily understood as being due to the fact that the
Sommerfeld enhancement has not saturated yet at the mass scale of the
top-level subhalo, such that it can be boosted up to it saturation
point by the contribution from the kinematically colder substructure.

@@

It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
predicted to be coming from the \dSphs\ is very large in a large class
of models. We have to stress here that our predicted fluxes are all
for the particle physics specific parameters given in Section
\ref{sec:partphys}, \ie, \sigmaannv\ = $3 \times 10^{-26}$ cm$^3$
s$^{-1}$, \mdm\ = 700 GeV and a particular form for the \DM\
annihilation spectrum. To predict the gamma-ray flux for a particular
model with different values of these particle physics parameters one
is required to calculate the particle physics parameter specific
factor to the \DM\ annihilation flux and appropriately rescale the
values given here. In order to compare these signals to the expected
background flux we will only consider contributions to the background
flux of extragalactic gamma rays. Although there are also
contributions to the background flux from pion-decay, inverse Compton
scattering, and Bremsstrahlung, at the high Galactic latitudes at
which the dwarf Spheroidals are found these are all of the same order
or smaller than the extragalactic background
\cite{2004ApJ...613..962S}. We have integrated the extragalactic gamma
ray flux as measured by EGRET over the energy range between 4 GeV --
250 GeV and multiplied it with the solid angle that the \dSphs\ span
on the sky. The extragalactic gamma ray flux is given by a power law
\cite{1998ApJ...494..523S}
\begin{equation}
\frac{\dd \Ngamma}{\dd E \,\dd A \, \dd t\,\dd \Omega} = k\,\left(\frac{E}{E_0}\right)^{-\alpha}\, ,
\end{equation}
where $k = 7.3 \times 10^{-6}$ cm$^{-2}$ s$^{-1}$ sr$^{-1}$
GeV$^{-1}$, $\alpha$ = 2.10, and $E_0 = 0.45$ GeV. We have included
the background flux of extragalactic photons in Figure
\ref{fig:dsph}. This background flux is in all cases of the order of
10$^{-9}$ -- 10$^{-10}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
annihilation boosted by substructure and Sommerfeld enhancement
outshine this background in most of the parameter space, and by a few
orders of magnitude in most models. Given that the flux above 100 MeV
over a 55 days time-span needed to separate a point source of \DM\
annihilation from the background by \Fermi\ to give a 5$\sigma$
detection is about 10$^{-8}$ photons cm$^{-1}$ s$^{-1}$ and that the
necessary fluxes scale roughly as $t_{\mbox{{\footnotesize
exp}}}^{-1/2}$ \cite{2008JCAP...07..013B}, the \DM\ annihilation of
these \dSphs\ should be detectable by \Fermi\ in many of the models
parametrized by ($\mv/\mdm,\alpha$). The Ursa Major II \dSph\ is the
only \dSph\ for which the \DM\ annihilation dominates the background
in all the regions of parameter space. This shows that the newly
discovered, most-dark-matter-dominated satellite galaxies are possibly
the best targets for the indirect detection of \DM. We conclude that
the prospects for detecting \DM\ annihilation from the \dSphs\ with
\Fermi\ are very good in models that include a significant Sommerfeld
enhancement.

@@

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_2.eps}
\caption{Left: Histogram of the predicted gamma-ray flux coming from
individual subhalos contained within the tidal radius of the main halo
in \VL\ II for three random observer placed at a distance of 8.5 kpc
from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
$\alpha = 10^{-3}$. Right: Histogram of the total boosts of the
subhalos for the same values of \mv/\mdm\ and $\alpha$.}%\
\label{fig:VL23_2}%
\end{figure}

@@

We only cover some typical cases for the Sommerfeld enhancement here
which suffices to show the general trend. We consider the particle
physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
smallest Sommerfeld enhancements in Figures \ref{fig:VL23_2} and
\ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
left panel of Figure \ref{fig:VL23_2} shows the distribution of
gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
right panel shows the distribution of the total boost factors. As can
be seen, the total boost factors are quite small: the typical boost
factor is only about 30 with the highest boost factors being about
100. The total gamma-ray flux coming from subhalos is then also quite
small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
the total gamma-ray flux is shown for 1,000 random observers at a
distance of 8.5 kpc from the galactic center. At most observation
locations this total flux is about 10$^{-7}$ cm$^2$ s$^{-1}$, with the
high-end tail not extending further than about half an order of
magnitude from this. For comparison the dashed histogram in Figure
\ref{fig:VL23_3} shows the same distribution of total gamma-ray flux
for different random observers in the absence of any Sommerfeld
enhancement. This confirms that the total gamma-ray flux with
Sommerfeld enhancement is about an order of magnitude larger than
without Sommerfeld enhancement.

@@

{\bf Acknowledgments:} It is a pleasure to thank Ilias Cholis, Mulin
Ding, David Hogg, and Neal Weiner for comments and assistance. This
research was partially supported by NASA (grant NNX08AJ48G).
